---
title: "ggplot2 review"
output: html_notebook
---


#1 First ggplot2 bar chart
```{r}
#Setup
library(ggplot2)
library(gcookbook)

head(climate)

csub<-subset(climate,Source=='Berkeley'&Year>=1900)
head(csub)
csub$pos<-csub$Anomaly10y>=0
csub

```


```{r}
ggplot(csub,aes(x=Year,y=Anomaly10y,fill=pos))+
         geom_bar(stat='identity',position='identity')
```

1 fill=
  pos is a column with value of T/F; fill=pos  is used for color
2 stat=
  geom_bar(stat='identity') the height represents y value, for discrete X and continuous Y
        or stat='bin'     the height of each bar equal to the number of cases in each group (histogram)

Position adjustments
3 position=
  position='identity'     seem to be useless
  position='dodge'         side by side 
  position='fill'      stack on top of one another, with normalized height(see proportion)
  position='stack'      stack on top of one another, with absolute height
  geom_point(position='jitter')   avoid overlapping

```{r}
ggplot(csub,aes(x=Year,y=Anomaly10y,fill=pos))+
  geom_bar(stat='identity',position='identity',color='black',size=0.25)+
  scale_fill_manual(values=c("#CCEEFF","#FFDDDD"),guide=FALSE)
```


color/size: the color/size of border
Scales
scale_fill_manual(values=
                  limits=
                  breaks=
                  names=  #title to use in legend
                  labels=  )
                  
guide=FALSE   get rid of legend

```{r}
ggplot(csub,aes(x=Year,y=Anomaly10y,fill=pos))+
  geom_bar(stat='identity',position='identity',color='black',size=0.25)+
  scale_fill_manual(values=c("#CCEEFF","#FFDDDD"),guide=FALSE)+
  labs(y="10 year Anomaly in degrees C")+
  ggtitle("Its getting hot in here!")+
  theme_minimal()

```

labels
ggtitle()
xlab()
ylab()
labs(title= , x= , y= )

theme
theme_bw()
theme_grey()
theme_classic()
theme_minimal()


#2 Intro to ggplot2
##Geometric objects and aesthetics
```{r}
#Basic setup
library(tidyverse)
setwd("C:/_Emory/2&4-675-Data Viz/In-class exercises/0910-Intro to ggplot2/Rgraphics/dataSets")
housing <- read_csv("landdata-states.csv")
head(housing[1:5])
```

```{r}
#histogram in ggplot2
ggplot(housing,aes(x=Home.Value))+
  geom_histogram()
```

```{r}
ggplot(filter(housing,State %in% c("MA","TX")),
       aes(x=Date,y=Home.Value,color=State))+
  geom_point()
```

using color instead of fill


points(scatterplots)
```{r}
hp2001Q1 <- filter(housing, Date == 2001.25) 
ggplot(hp2001Q1,
       aes(y = Structure.Cost, x = Land.Value)) +
  geom_point()

ggplot(hp2001Q1,
       aes(y = Structure.Cost, x = log(Land.Value))) +
  geom_point()
```

prediction line
```{r}
hp2001Q1$pred.SC <- predict(lm(Structure.Cost ~ log(Land.Value), data = hp2001Q1))

p1 <- ggplot(hp2001Q1, aes(x = log(Land.Value), y = Structure.Cost))

p1 + geom_point(aes(color = Home.Value)) +
  geom_line(aes(y = pred.SC))
```

smoothers
```{r}
p1 +
  geom_point(aes(color = Home.Value)) +
  geom_smooth()
```

text
```{r}
p1 + 
  geom_text(aes(label=State), size = 3)

library("ggrepel")
p1 + 
  geom_text_repel(aes(label=State), size = 3)

p1 + 
  geom_point() + 
  geom_text_repel(aes(label=State), size = 3)
```

geom_text(aes(label=))
geom_text_repel: avoid overlapping

```{r}
p1 +
  geom_point(aes(color=Home.Value, shape = region))
```


Exercise I
```{r}
setwd("C:/_Emory/2&4-675-Data Viz/In-class exercises/0910-Intro to ggplot2/Rgraphics/dataSets")
dat <- read_csv("EconomistData.csv")
dat
```

These data consist of Human Development Index and Corruption Perception Index scores for several countries.

1 Create a scatter plot with CPI on the x axis and HDI on the y axis.
2 Color the points blue.
3 Map the color of the the points to Region.
4 Make the points bigger by setting size to 2
5 Map the size of the points to HDI.Rank

```{r}
ggplot(dat,aes(x=CPI,y=HDI))+
  geom_point(color="blue",size=2)
  
ggplot(dat,aes(x=CPI,y=HDI))+
  geom_point(aes(color=Region),size=2)

ggplot(dat,aes(x=CPI,y=HDI,color=Region))+
  geom_point()

ggplot(dat,aes(x=CPI,y=HDI))+
  geom_point(aes(color=Region,size=HDI.Rank))

```

color="bule"  quotation mark

color/size are included in aes

##Statistical transformations


```{r}
p2 <- ggplot(housing, aes(x = Home.Value))
p2 + geom_histogram()

p2 + geom_histogram(stat = "bin", binwidth=4000)
```


```{r}
housing.sum <- aggregate(housing["Home.Value"], housing["State"], FUN=mean)
rbind(head(housing.sum), tail(housing.sum))

#ggplot(housing.sum, aes(x=State, y=Home.Value)) + 
  #geom_bar()

ggplot(housing.sum, aes(x=State, y=Home.Value)) + 
  geom_bar(stat="identity")
```

Exercise II
1 Re-create a scatter plot with CPI on the x axis and HDI on the y axis (as you did in the previous exercise).
2 Overlay a smoothing line on top of the scatter plot using geom_smooth.
3 Overlay a smoothing line on top of the scatter plot using geom_smooth, but use a linear model for the predictions. Hint: see ?stat_smooth.
4 Overlay a smoothing line on top of the scatter plot using geom_line. Hint: change the statistical transformation.
5 BONUS: Overlay a smoothing line on top of the scatter plot using the default loess method, but make it less smooth. Hint: see ?loess.


```{r}
ggplot(dat, aes(x = CPI, y = HDI)) +
  geom_point()+
  geom_smooth(method="lm")

ggplot(dat, aes(x = CPI, y = HDI)) +
  geom_point()+
  geom_line(stat="smooth",method="loess")

ggplot(dat, aes(x = CPI, y = HDI)) +
  geom_point()+
  geom_smooth(method="loess",span=0.4)
```

geom_smooth(method="lm")
geom_line(stat="smooth",method="loess")

##Scales

```{r}
p3 <- ggplot(housing,aes(x = State, y = Home.Price.Index)) + 
        theme(legend.position="top", axis.text=element_text(size = 6))
p4 <- p3 + geom_point(aes(color = Date),
              alpha = 0.5,size = 1.5,
              position = position_jitter(width = 0.25, height = 0))
p4
```

size of x/y axis tickers: axis.text=element_text(size = 6)

```{r}
p4 + scale_x_discrete(name="State Abbreviation") +
  scale_color_continuous(name="",
                         breaks = c(1976, 1994, 2013),
                         labels = c("'76", "'94", "'13"))
```


change xlabs: scale_x_discrete(name="State Abbreviation")

```{r}
p4 +
  scale_x_discrete(name="State Abbreviation") +
  scale_color_continuous(name="",
                         breaks = c(1976, 1994, 2013),
                         labels = c("'76", "'94", "'13"),
                         low = "blue", high = "red")

#Another Choice
#                       low = muted("blue"), high = muted("red"))
```

```{r}
library(scales)

p4 +
  scale_color_gradient2(name="",
                        breaks = c(1976, 1994, 2013),
                        labels = c("'76", "'94", "'13"),
                        low = muted("blue"),
                        high = muted("red"),
                        mid = "gray60",
                        midpoint = 1994)
```


Exercise III
1 Create a scatter plot with CPI on the x axis and HDI on the y axis. Color the points to indicate region.
2 Modify the x, y, and color scales so that they have more easily-understood names (e.g., spell out “Human development Index” instead of “HDI”).
3 Modify the color scale to use specific values of your choosing. Hint: see ?scale_color_manual.


there may exist some mistakes//OSOSOSOSOS
```{r}
ggplot(dat,aes(x=CPI,y=HDI))+
  geom_point(aes(color=Region))


ggplot(dat,aes(x=CPI,y=HDI))+
  geom_point(aes(color=Region))+
  scale_x_discrete(name="Corruption Perception Index")+
  scale_y_discrete(name="Human development Index")+
  scale_color_discrete(name="Region of the world")

ggplot(dat,aes(x=CPI,y=HDI))+
  geom_point(aes(color=Region))+
  labs(x="Corruption Perception Index",y="Human development Index")+
  scale_color_discrete(name="Region of the world")

ggplot(dat, aes(x = CPI, y = HDI)) +
geom_point(aes(color = "Region")) +
scale_x_continuous(name = "Corruption Perception Index") +
scale_y_continuous(name = "Human Development Index") +
  scale_color_manual(name = "Region of the world",
                     values = c("#24576D",
                                "#099DD7",
                                "#28AADC",
                                "#248E84",
                                "#F2583F",
                                "#96503F"))


```

##Faceting

faceting: make more plots instead of put all in one plot

```{r}
p5 <- ggplot(housing, aes(x = Date, y = Home.Value))
p5 + geom_line(aes(color = State))  

p5<-p5 + geom_line() +
   facet_wrap(~State, ncol = 10)
p5
```

Use ~State for identify different charts
facet_wrap(~State, ncol = 10)

##Themes

```{r}
p5 + theme_linedraw()
p5 + theme_light()
p5 + theme_minimal() +
  theme(text = element_text(color = "turquoise"))
```

Create a theme you like
```{r}
theme_new <- theme_bw() +
  theme(plot.background = element_rect(size = 1, color = "blue", fill = "black"),
        text=element_text(size = 12, family = "Serif", color = "ivory"),
        axis.text.y = element_text(colour = "purple"),
        axis.text.x = element_text(colour = "red"),
        panel.background = element_rect(fill = "pink"),
        strip.background = element_rect(fill = muted("orange")))

p5 + theme_new
```

##Challenge solution

```{r}
pc1 <- ggplot(dat, aes(x = CPI, y = HDI, color = Region))
pc1 + geom_point()
```


To complete this graph we need to:

[ ] add a trend line
[ ] change the point shape to open circle
[ ] change the order and labels of Region
[ ] label select points
[ ] fix up the tick marks and labels
[ ] move color legend to the top
[ ] title, label axes, remove legend title
[ ] theme the graph with no vertical guides
[ ] add model R2 (hard)
[ ] add sources note (hard)
[ ] final touches to make it perfect (use image editor for this)

```{r}
pc2 <- pc1 +
  geom_smooth(mapping = aes(linetype = "r2"),
              method = "lm",
              formula = y ~ x + log(x), se = FALSE,
              color = "red")
pc2 + geom_point()
```

geom_smooth()
se=FALSE : do not show the  range

```{r}
pc2 +
  geom_point(shape = 1, size = 4)

pc3 <- pc2 + geom_point(shape = 1, size = 2.5, stroke = 1.25)
pc3
```

stroke: the width of the border


labelling points
```{r}
pointsToLabel <- c("Russia", "Venezuela", "Iraq", "Myanmar", "Sudan",
                   "Afghanistan", "Congo", "Greece", "Argentina", "Brazil",
                   "India", "Italy", "China", "South Africa", "Spane",
                   "Botswana", "Cape Verde", "Bhutan", "Rwanda", "France",
                   "United States", "Germany", "Britain", "Barbados", "Norway", "Japan",
                   "New Zealand", "Singapore")
(pc4 <- pc3 +
  geom_text(aes(label = Country),
            color = "gray20",
            data = filter(dat, Country %in% pointsToLabel)))

```



#3 EDA

##ECDF

```{r}
#Data Preparation
set.seed(1234)
df<-data.frame(height=round(rnorm(200,mean=60,sd=15)))
height.q<-quantile(df$height)
height.q
```


Plot1
```{r}
ggplot(df,aes(height))+
  stat_ecdf(geom='step')+
  labs(title='Empirical Cumulative Distribution Function',y='F(height)',x='Height in inches')+
  theme_classic()+
  geom_vline(aes(xintercept=quantile(height)[2]),linetype='dashed')+# first quartile
  geom_vline(aes(xintercept=quantile(height)[4]),linetype='dashed')+
  geom_hline(yintercept=0.25,linetype='dashed')+
  geom_hline(yintercept=0.75,linetype='dashed')
```

ecdf:
stat_ecdf(geom="step")
ref lines:
vertical:geom_vline() using variables: height, thus need aes(xintercept=quantile(height)[2])
horizontal:geom_hline() yintercept=0.25

subset data for add specific ref line
geom_vline(data=subset(tophitters2001,lg=='AL'),aes(xintercept=quantile(avg)[3]))

linetype='dashed'


Plot2
```{r}
#First two lines generate the same plot

ggplot(tophitters2001,aes(avg,fill=lg,color=lg))+
  stat_ecdf(geom='step')

ggplot(tophitters2001,aes(avg))+
  stat_ecdf(geom='step',aes(color=lg))+
  geom_hline(yintercept=0.5,linetype='dashed')+
  geom_vline(data=subset(tophitters2001,lg=='AL'),aes(xintercept=quantile(avg)[3]),
             linetype='dashed')+
  geom_vline(data=subset(tophitters2001,lg=='NL'),aes(xintercept=quantile(avg)[3]),
             linetype='dashed')+
  labs(title='MLB Top Hitters 2001 Batting Average',x='Batting Average',y='ECDF')+
  theme_classic()
```

##Histogram/violin/beeswarm showing spread/distribution
```{r}
#library(gcookbook)
View(tophitters2001)

ggplot(tophitters2001,aes(avg,fill=lg))+
  geom_histogram(alpha=0.5)

ggplot(tophitters2001,aes(lg,avg))+
  geom_violin(aes(fill=lg))+
  geom_jitter(width=0.1)

library(ggbeeswarm)

ggplot(tophitters2001,aes(lg,avg))+
  geom_beeswarm(aes(color=lg))
```

##Histogram

```{r}
#Data preparation
library(fGarch)
set.seed(1953)

norm <- rnorm(1000, 50, 10)

norm.highMean <- rnorm(1000, 75, 10)

norm.highSD <- rnorm(1000, 50, 30)

#xi:skewness parameter
norm.NegSkew <- rsnorm(1000, mean=50, sd=10, xi=-10)

norm.PosSkew <- rsnorm(1000, mean=50, s=10, xi=10)

```

```{r}
#par(mfrow=c(5,1))
hist(norm, main="Norm(50,10)", xlim=c(-100,200), breaks=10)
hist(norm.highMean, main="Norm(75,10)", xlim=c(-100,200), breaks=10)
hist(norm.highSD, main="Norm(50,30)", xlim=c(-100,200), breaks=10)
hist(norm.NegSkew, main="Norm(50, 10, -10)", xlim=c(-100,200), breaks=10)
hist(norm.PosSkew, main="Norm(50, 10, 10)", xlim=c(-100,200), breaks=10)
```

breaks: number of bins


##QQplot

```{r}
qqplot(norm, norm, main="Q-Q Plot Nor(50,10) vs Norm(50,10")
abline(0, 1, col="red")

qqplot(norm, norm.highMean, main="Q-Q Plot Norm(50,10) vs Norm(75,10)")
abline(0, 1, col="red")

qqplot(norm, norm.highSD, main="Q-Q Plot Norm(50,10) vs Norm(50,30)")
abline(0, 1, col="red")

qqplot(norm, norm.NegSkew, main="Q-Q Plot Norm(50,10) vs Norm(50,10,-10)")
abline(0, 1, col="red")

qqplot(norm, norm.PosSkew, main="Q-Q Plot Norm(50,10) vs Norm(50,10,+10)")
abline(0, 1, col="red")
```

abline(a,b,col="red")
color doesn't work, must be col
y=bx+a
intercept: a  slope: b


QQ plot for comparing two distribution:Iris example
```{r}
setosa <- subset(iris, Species=="setosa")
hist(setosa$Sepal.Length, xlim=c(1,10))
versicolor <- subset(iris, Species=="versicolor")
hist(versicolor$Sepal.Length, xlim=c(1,10))
qqplot(setosa$Sepal.Length, versicolor$Sepal.Length, xlim=c(1,10), ylim=c(1,10))
abline(0,1, col="red")
```


##Scatter plots
```{r}
data(oly12, package="VGAMdata")
ggplot(oly12, aes(Height, Weight)) + geom_jitter() +
  ggtitle("Athletes at the London Olympics 2012") + theme_minimal()
```


```{r}
data(geyser, package="MASS")
ggplot(geyser, aes(duration, waiting)) + geom_point() +
  geom_density2d() + 
  ggtitle("Old Faithful Eruption Waiting Time vs Duration (Minutes)") +
  theme_minimal()
```

drawing contours: geom_density2d() 


```{r}
oly12S <- within(oly12, Sport <- abbreviate(Sport, 12))
ggplot(oly12S, aes(Height, Weight)) +
  geom_point(size=1, alpha=0.5) + 
  facet_wrap(~Sport)+
  ggtitle("Weight and Height by Sport") +
  theme_minimal()
```

In GGally

ggpairs: showing scatterplots, distribution, and correlation [correlation matrix]
```{r}
data(crime.us, package="VGAMdata")
crime.usR <- crime.us

names(crime.usR) <- gsub("*Rate", "", names(crime.usR))
names(crime.usR)[19:20] <- c("Larceny", "MotorVTheft")
library(GGally)
ggpairs(crime.usR[, c(13:16,18:20)],
        title="Crime rates in the USA",
        diag=list(continuous='density'), axisLabels = 'none')
```


In ggExtra

rug
```{r}
library(ggExtra)

p <- ggplot(mtcars, aes(wt, mpg)) +
  geom_point()
p
p + geom_rug()
p + geom_rug(sides="b")    # Rug on bottom only
p + geom_rug(sides="trbl") # All four sides

# Use jittering to avoid overplotting for smaller datasets
ggplot(mpg, aes(displ, cty)) +
  geom_point() +
  geom_rug()

ggplot(mpg, aes(displ, cty)) +
  geom_jitter() +
  geom_rug(alpha = 1/2, position = "jitter", colour="red") +
  theme_classic()
```

Another way to see distribution, just see it on axis
geom_rug(sides="b") bottom
geom_rug(alpha = 1/2, position = "jitter", colour="red")


ggMarginal: add reference distributions
```{r}
histP <- ggplot(mpg, aes(displ, cty)) +
  geom_jitter() +
  theme_classic()
ggExtra::ggMarginal(histP, type="histogram", fill="red")

ggMarginal(histP, type="histogram", fill="red")
#?ggMarginal
ggMarginal(histP, type="density", fill="blue")
ggMarginal(histP, type="densigram", fill="red")
```


Other histogram example

```{r}
library(ggthemes)

birth<-read.csv('http://datasets.flowingdata.com/birth-rate.csv')
hist(birth$X2008,breaks = 20)
```


```{r}
ggplot(birth,aes(birth$X2008))+
  geom_histogram(binwidth=5,col='red',fill='powderblue',alpha=0.2)+
  labs(title = 'Global Distributions of Birth Rates',
       x='Live Births per 1,000',
       y='County Count')+
  ggthemes::theme_tufte()

```

geom_histogram(binwidth=5,col='red',fill='powderblue',alpha=0.2)
binwidth: each bin's width
col: outline color
fill: inside color

```{r}
ggplot(faithful,aes(x=waiting))+geom_histogram()
ggplot(faithful,aes(x=waiting))+geom_density()
ggplot(faithful,aes(x=waiting))+geom_line(stat='density')+
  expand_limits(y=0)
ggplot(faithful,aes(x=waiting))+geom_density(fill='blue',color=NA,#color=NA remove the outline first
                                             alpha=0.2)+
  geom_line(stat='density')+#add a new outline using geom_line
  xlim(35,105)
```

geom_histogram()
geom_density() <=> geom_line(stat='density')

```{r}
library(MASS)
birthwt1<-birthwt

birthwt1$smoke<-factor(birthwt1$smoke)

ggplot(birthwt1,aes(x=bwt,fill=smoke))+
  geom_density(alpha=0.3)+
  labs(title='Birth Weight and Smoking',x='Birth Weight(in grams)',
       y='Density',fill='Mother Smoked')+#fill-mother smoked as legend titile
  scale_fill_manual(labels=c('No','Yes'),values=c('navy','maroon'))+#rewrite legend, dont use legend as variable names 
  theme_few()

ggplot(birthwt1,aes(x=bwt,fill=smoke))+
  geom_histogram(alpha=0.3,position='identity')+
  labs(title='Birth Weight and Smoking',x='Birth Weight(in grams)',
       y='Density',fill='Mother Smoked')+#fill-mother smoked as legend titile
  scale_fill_manual(labels=c('No','Yes'),values=c('navy','maroon'))+#rewrite legend, dont use legend as variable names 
  theme_few()

ggplot(birthwt1,aes(x=bwt,fill=smoke))+
  geom_histogram(alpha=0.3,position='dodge')+
  labs(title='Birth Weight and Smoking',x='Birth Weight(in grams)',
       y='Density',fill='Mother Smoked')+#fill-mother smoked as legend titile
  scale_fill_manual(labels=c('No','Yes'),values=c('navy','maroon'))+#rewrite legend, dont use legend as variable names 
  theme_few()

ggplot(birthwt1,aes(x=bwt,fill=smoke))+
  geom_histogram(alpha=0.3,position='stack')+
  labs(title='Birth Weight and Smoking',x='Birth Weight(in grams)',
       y='Density',fill='Mother Smoked')+#fill-mother smoked as legend titile
  scale_fill_manual(labels=c('No','Yes'),values=c('navy','maroon'))+#rewrite legend, dont use legend as variable names 
  theme_few()
```

position= identity/dodge/stack
setting legends and axis
scale_fill_manual()


#4 Homework 1

Plot1

```{r}
library(readxl)

library(scales)
library(psych)
library(ggplot2)
library(ggthemes)
library(dplyr)
library(grid)
library(showtext)
library(Cairo)
library(readr)

```


```{r}
HBR_CEO_data <- read_excel("C:/_Emory/2&4-675-Data Viz/Assignments/#1-Due Oct 2 Wed/HBR-CEO-data-PV.xlsx")

data<-HBR_CEO_data

# Bar chart and flip the axis
p11<-ggplot(data,aes(x=reorder(data$`CEO NAME`,-data$Rank),data$Pay,fill=data$Category))+
  geom_bar(stat="identity",position='dodge')+
  scale_y_continuous(labels = comma,breaks=seq(0,50000,5000))+
  coord_flip()

# Add themes, title, legends and other annotations
p12<-p11+theme_economist()+
  labs(title='Top 10 US CEO Compensation',x='',y='USD in thousand')+
  theme(legend.position=c(0.91,0.947),legend.title=element_blank(),
        plot.title = element_text(hjust = 0.5))

p12

```


Plot2

```{r}
BigMac <- read_csv("C:/_Emory/2&4-675-Data Viz/Assignments/#1-Due Oct 2 Wed/BigMac.csv", 
                   col_types = cols(date = col_date(format = "%Y/%m/%d")))


# Line, point, setting axis and ticks
p21<-ggplot(BigMac,aes(x=BigMac$date,y=BigMac$FX,
                      group=BigMac$Category,fill=BigMac$Category,
                      color=BigMac$Category))+
  coord_cartesian(ylim=c(0, 8.5))+scale_y_continuous(breaks=seq(0,9,0.5))+
  geom_point(size=1.5)+
  geom_line(size=0.8)
  

# Titles, themes
p22<-p21+theme_stata()+
  labs(title='CNY/USD Exchange Rate',x='',y='CNY per USD')+
  theme(legend.position=c(0.9,0.92),legend.title=element_blank(),
        plot.title = element_text(face='bold',hjust = 0.5))


# Annotations
p23<-p22+annotate("text", x=BigMac$date[6]+53, y=7.2, label="2005-07-21 \n Abolished Yuan-Dollar peg \n Adopted floating rate",fontface='bold')+
  annotate("text", x=BigMac$date[24]+40, y=5.5, label="2015-08-11 \n Improved central parity system \n to better reflect market development",fontface='bold')

# Reference line
p24<-p23+geom_vline(xintercept=BigMac$date[6]+53,linetype=2,size=.25,colour="grey")+
  geom_vline(xintercept=BigMac$date[24]+40,linetype=2,size=.25,colour="grey")
  
# Annotations for Stages
  
p25<-p24+annotate("text", x=BigMac$date[3]+60, y=1.5, label="Stage 1",fontface='bold')+
  annotate("text", x=BigMac$date[15]-30, y=1.5, label="Stage 2",fontface='bold')+
  annotate("text", x=BigMac$date[28]+40, y=1.5, label="Stage 3",fontface='bold')

p25
```



Others
geom_step()














